Supercritical fluid extraction of essential oil from chamomile flowers: modelling, and parameter estimation

### 2024-03-13 SFE_article_Parameter_Estimation_short - Chamomile/Sections/Model.tex
13.03

			{\footnotesize
				\begin{equation} \label{Model_kinetic_basic}
					{\color{black}r_e}(t,z) = \cfrac{{\color{black}D_i}({\color{black}T}(t,z), {\color{black}P}(t))}{{\color{black} \mu l^2} }\left({\color{black}c_s}(t,z) - {\color{black}c_P^*}(t,z) \right)
			\end{equation} }

@@

			{\footnotesize
				\begin{equation}
					\label{Model_kinetic_no_sat}
					{\color{black}r_e}(t,z) = -\cfrac{{\color{black}D_i}({\color{black}T}(t,z), {\color{black}P}(t))}{{\color{black} \mu l^2} }\left({\color{black}c_s}(t,z) - \cfrac{{\color{black}\rho_s}}{{\color{black}k_m}({\color{black}T}(t,z)){\color{black}\rho_f}({\color{black}T}(t,z),{\color{black}P}(t))}  {\color{black}c_f}(t,z) \right)
			\end{equation} }

@@

		\subsubsection{Heat balance} \label{CH: heat_balance}

		\iffalse
		 {\color{black}
			The heat balance (Equation \ref{Model_heat}) consists of the convective and diffusive terms. It follows the assumption of a pseudo-homogeneous phase, which properties are the mean between fluid and solid phases. We consider no heat loss through the wall, and there is no heat generation in the system. Therefore, the temperature of the extractor can be changed only by manipulating the temperature of the inlet stream %$T_{In}(t)$.

@@

%		\todo[inline]{Draft of new heat equation starts here. TODO: \\ Check if the heat cpacity of solid phase needs to be added \\ Elaborate why pressure is know}
		\fi

@@

		\subsubsection{Initial and boundary conditions} 
		It is assumed that the solvent is free of solute at the entrance of the extractor and that all the solid particles have the same initial solute content ${\color{black}c_{s0}}$. As the residence time is much shorter than the sampling time, the initial state estimate for the concentration of the solute in the fluid phase would be not reliable. Considering that it is assumed that the $c_{f0}=0$. Moreover, it is considered that the initial temperature of the extractor in every place is the same and described by ${\color{black}h_0}$. Therefore, the initial conditions employed in the simulation are:

@@

		\subsubsection{State-space representation} \label{CH: State_space}

@@

		Function ${\color{black}\bar{\phi}}$ are transformed to a corresponding set of $N_z$ discretized equations denoted as ${\color{black}G}$. The state-space model ${\color{black}\chi}(t,z)$ after the discretization is represented by $\dot{{\color{black}x}}(t)$.

			where ${\color{black}x} \in \mathbb{R}^{N_x = 3N_z} $ and ${\color{black}\Theta} \in \mathbb{R}^{N_\Theta =  N_{\theta} + N_u } $, $N_{\theta}$ is the number of model parameters, $N_{u}$ is the number of control variables.

@@

			The method of lines is used to transform the process model equations into a set of ODEs denoted as ${\color{black}G}({\color{black}x}(t);{\color{black}\Theta})$. The partial derivatives in $z$-direction are computed using a first-order and second-order finite difference approximation. The backward finite difference is used to approximate the first-order derivative, while the central difference scheme is used to approximate the second-order derivative. The length of the fixed bed is divided into $N_z$ equally distributed points in $z$-direction.

>>>>

			{\footnotesize
				\begin{equation} \label{Model_kinetic_basic}
					{\color{black}r_e}(t,z) = \cfrac{{\color{black}D_i}({\color{black}T}(t,z), {\color{black}P}(t), F(t))}{{\color{black} \mu l^2} }\left({\color{black}c_s}(t,z) - {\color{black}c_P^*}(t,z) \right)
			\end{equation} }

@@

			{\footnotesize
				\begin{equation}
					\label{Model_kinetic_no_sat}
					{\color{black}r_e}(t,z) = -\cfrac{{\color{black}D_i}({\color{black}T}(t,z), {\color{black}P}(t), F(t))}{{\color{black} \mu l^2} }\left({\color{black}c_s}(t,z) - \cfrac{{\color{black}\rho_s}c_f(t,z)}{{\color{black}k_m}({\color{black}T}(t,z)){\color{black}\rho_f}({\color{black}T}(t,z),{\color{black}P}(t))}   \right)
			\end{equation} }

@@

		\iffalse	
		\subsubsection{Heat balance} \label{CH: heat_balance}

		 {\color{black}
			The heat balance (Equation \ref{Model_heat}) consists of the convective and diffusive terms. It follows the assumption of a pseudo-homogeneous phase, which properties are the mean between fluid and solid phases. We consider no heat loss through the wall, and there is no heat generation in the system. Therefore, the temperature of the extractor can be changed only by manipulating the temperature of the inlet stream %$T_{In}(t)$.

@@

%		\todo[inline]{Draft of new heat equation starts here. TODO: \\ Check if the heat cpacity of solid phase needs to be added \\ Elaborate why pressure is know}

@@

		\subsubsection{Initial and boundary conditions} 
		It is assumed that the solvent is free of solute at the entrance of the extractor and that all the solid particles have the same initial solute content ${\color{black}c_{s0}}$. As the residence time is much shorter than the sampling time, the initial state estimate for the concentration of the solute in the fluid phase would be not reliable. Considering that it is assumed that the $c_{f0}=0$. Moreover, it is considered that the extractor is isothermal and described by ${\color{black}h_0}$. Therefore, the initial conditions employed in the simulation are:

@@

		The method of lines is used to transform the process model equations into a set of ODEs denoted as ${\color{black}G}({\color{black}x}(t);{\color{black}\Theta})$. The partial derivatives in $z$-direction are computed using a first-order and second-order finite difference approximation. The backward finite difference is used to approximate the first-order derivative, while the central difference scheme is used to approximate the second-order derivative. The length of the fixed bed is divided into $N_z$ equally distributed points in $z$-direction. The state-space model after the discretization is represented by $\dot{{\color{black}x}}(t)$.

		where ${\color{black}x} \in \mathbb{R}^{N_x = 3N_z} $ and ${\color{black}\Theta} \in \mathbb{R}^{N_\Theta =  N_{\theta} + N_u } $, $N_{\theta}$ is the number of parameters, $N_{u}$ is the number of control variables.

		\iffalse
		\subsubsection{State-space representation} \label{CH: State_space}

@@

		\fi

		%Function ${\color{black}\bar{\phi}}$ are transformed to a corresponding set of $N_z$ discretized equations denoted as ${\color{black}G}$.
		%The state-space model ${\color{black}\chi}(t,z)$ after the discretization is represented by $\dot{{\color{black}x}}(t)$.

@@

### 2024-03-13 SFE_article_Parameter_Estimation_short - Chamomile/Sections/Results_Chamomile.tex
13.03

	\begin{figure*}[!h]
		\centering
		\begin{subfigure}{0.24\textwidth}
			\centering
			\includegraphics[trim = 0.0cm 0.0cm 0.0cm 0.0cm,clip, width=\textwidth]{/Results_estimation/Correlation_Di_Re.png}
			\caption{Linear regression $D_i^R = f(Re)$}
			\label{fig: Correlations_Di_Re}
		\end{subfigure}
		\hfill
		\begin{subfigure}{0.24\textwidth}
			\centering
			\includegraphics[trim = 0.0cm 0.0cm 0.0cm 0.0cm,clip, width=\textwidth]{/Results_estimation/Correlation_Gamma_Re.png}
			\caption{Linear regression $\Upsilon = f(Re)$}
			\label{fig: Correlations_Gamma_Re}
		\end{subfigure}
		\hfill
		\begin{subfigure}{0.24\textwidth}
			\centering
			\includegraphics[trim = 0.0cm 0.0cm 0.0cm 0.0cm,clip, width=\textwidth]{/Results_estimation/Di_Re_F.png}
			\caption{Multiple linear regression $D_i^R = f(Re, F)$}
			\label{fig: Correlations_Di_Re_F}
		\end{subfigure}
		\hfill
		\begin{subfigure}{0.24\textwidth}
			\centering
			\includegraphics[trim = 0.0cm 0.0cm 0.0cm 0.0cm,clip, width=\textwidth]{/Results_estimation/Gamma_Re_F.png}
			\caption{Multiple linear regression $\Upsilon = f(Re, F)$}
			\label{fig: Correlations_Gamma_Re_F}
		\end{subfigure}
		\caption{Parameter estimation results}
		%\label{fig: Correlations_surface}
	\end{figure*}

@@

	\fi	
	The parameter estimation results are combined to analyse the relationship between the obtained parameters and the operating conditions. Unlike traditional methods that employ a combination of Reynolds, Schmidt, and Sherwood numbers to find correlations—omitted here due to the insignificance of axial diffusion—the approach in this study leverages the Reynolds number $\left(Re = \frac{(2r) \cdot \rho_f \cdot u}{\mu}\right)$ as the sole independent variable. Using the Reynolds number has the advantage of considering the influence of all the control variables (temperature, pressure and flow rate), which means it can be uniquely defined by selecting operating conditions.

@@

	\begin{figure*}[!h]
		\centering
		\begin{subfigure}{0.3\textwidth}
			\centering
			\includegraphics[trim = 0.0cm 0.0cm 0.0cm 0.0cm,clip, width=\textwidth]{/Results_estimation/Fit_Di_Gamma_1_4_correlation.png}
			\caption{Results of parameter estimation for experiments at $6.67\times 10^{-5}$ [kg/s] and temperature of 40 $[^\circ C]$}
			\label{fig: Fit_1_4_Di_Gamma_correlation}
		\end{subfigure}
		\hfill
		\begin{subfigure}{0.3\textwidth}
			\centering
			\includegraphics[trim = 0.0cm 0.0cm 0.0cm 0.0cm,clip, width=\textwidth]{/Results_estimation/Fit_Di_Gamma_5_8_correlation.png}
			\caption{Results of parameter estimation for experiments at $6.67\times 10^{-5}$ [kg/s] and temperature of 30 $[^\circ C]$}
			\label{fig: Fit_5_8_Di_Gamma_correlation}
		\end{subfigure}
		\hfill
		\begin{subfigure}{0.3\textwidth}
			\centering
			\includegraphics[trim = 0.0cm 0.0cm 0.0cm 0.0cm,clip, width=\textwidth]{/Results_estimation/Fit_Di_Gamma_9_12_correlation.png}
			\caption{Results of parameter estimation for experiments at $3.33\times 10^{-5}$ [kg/s]}
			\label{fig: Fit_9_12_Di_Gamma_correlation}
		\end{subfigure}
		\caption{Parameter estimation results}
		\label{fig: Fit_Di_Gamma_correlation}
	\end{figure*}

>>>>

@@

	\fi

	The parameter estimation results are combined to analyse the relationship between the obtained parameters and the operating conditions. Unlike traditional methods that employ a combination of Reynolds, Schmidt, and Sherwood numbers to find correlations—omitted here due to the insignificance of axial diffusion—the approach in this study leverages the Reynolds number $\left(Re = \frac{(2r) \cdot \rho_f \cdot u}{\mu}\right)$ as the sole independent variable. Using the Reynolds number has the advantage of considering the influence of all the control variables (temperature, pressure and flow rate), which means it can be uniquely defined by selecting operating conditions.

	\begin{figure}[!h]
		\centering
		\begin{subfigure}{0.48\columnwidth}
			\centering
			\includegraphics[trim = 0.0cm 0.0cm 0.0cm 0.0cm,clip, width=\columnwidth]{/Results_estimation/Correlation_Di_Re.png}
			\caption{Linear regression $D_i^R = f(Re)$}
			\label{fig: Correlations_Di_Re}
		\end{subfigure}
		\hfill
		\begin{subfigure}{0.48\columnwidth}
			\centering
			\includegraphics[trim = 0.0cm 0.0cm 0.0cm 0.0cm,clip, width=\columnwidth]{/Results_estimation/Correlation_Gamma_Re.png}
			\caption{Linear regression $\Upsilon = f(Re)$}
			\label{fig: Correlations_Gamma_Re}
		\end{subfigure}
		\hfill
		\begin{subfigure}{0.48\columnwidth}
			\centering
			\includegraphics[trim = 0.0cm 0.0cm 0.0cm 0.0cm,clip, width=\columnwidth]{/Results_estimation/Di_Re_F.png}
			\caption{Multiple linear regression $D_i^R = f(Re, F)$}
			\label{fig: Correlations_Di_Re_F}
		\end{subfigure}
		\hfill
		\begin{subfigure}{0.48\columnwidth}
			\centering
			\includegraphics[trim = 0.0cm 0.0cm 0.0cm 0.0cm,clip, width=\columnwidth]{/Results_estimation/Gamma_Re_F.png}
			\caption{Multiple linear regression $\Upsilon = f(Re, F)$}
			\label{fig: Correlations_Gamma_Re_F}
		\end{subfigure}
		\caption{Parameter estimation results}
		%\label{fig: Correlations_surface}
	\end{figure}

@@

### 2024-03-13 SFE_article_Parameter_Estimation_short - Chamomile/Sections/Saturation.tex
13.03

	{\scriptsize
		\begin{equation}
			\label{Model_kinetic}
				{\color{black}r_e}(t,z) = -\cfrac{{\color{black}D_i^R}({\color{black}T}(t,z), {\color{black}P}(t)) {\color{black} \gamma}({\color{black} c_s} (t,z)) }{{\color{black} \mu l^2} } \left({\color{black}{\color{black} c_s} }(t,z)  - \cfrac{{\color{black}\rho_s}}{{\color{black}k_m}({\color{black}T}(t,z)){\color{black}\rho}({\color{black}T}(t,z),{\color{black}P}(t))}  {\color{black}c_f}(t,z) \right)
		\end{equation} }

>>>>

	{\scriptsize
		\begin{equation}
			\label{Model_kinetic}
				{\color{black}r_e}(t,z) = -\cfrac{{\color{black}D_i^R}({\color{black}T}(t,z), {\color{black}P}(t), F(t)) {\color{black} \gamma}({\color{black} c_s} (t,z)) }{{\color{black} \mu l^2} } \left({\color{black}{\color{black} c_s} }(t,z)  - \cfrac{{\color{black}\rho_s}c_f(t,z)}{{\color{black}k_m}({\color{black}T}(t,z)){\color{black}\rho}({\color{black}T}(t,z),{\color{black}P}(t))}  \right)
		\end{equation} }

### 2024-03-13 SFE_article_Parameter_Estimation_short - Chamomile/Sections/introduction_imp.tex
13.03

	This study investigates the extraction of essential oil from chamomile flowers (Matricaria chamomilla L.) via supercritical fluid extraction techniques and modelling of this process. Chamomile is a medicinal herb widely cultivated in southern and eastern Europe—such as Germany, Hungary, France, and Russia. It can be found outside of Europe in Brazil as discussed by \citet{Singh2011}. This plant is distinguished by its hollow, bright gold cones, housing disc or tubular florets and surrounded by about fifteen white ray or ligulate florets. Chamomile has been used for its medicinal benefits, serving as an anti-inflammatory, antioxidant, mild astringent, and healing remedy. Chamomile's aqueous extract is widely used as a gentle sedative to calm nerves and mitigate anxiety, hysteria, nightmares, insomnia, and other sleep-related conditions, according to \citet{Srivastava2009}. \citet{Orav2010} reported that oil yields from dried chamomile samples ranged from 0.7 to 6.7 mL kg$^{-1}$. Notably, the highest yields of essential oil, between 6.1 and 6.7 mL kg$^{-1}$, were derived from chamomile sourced from Latvia and Ukraine, while chamomile from Armenia exhibited a lower oil content of 0.7 mL kg$^{-1}$.

@@

	The organization of the study is as follows: Chapter \ref{CH: Thermodynamic} offers an overview of supercritical fluids, aiming to acquaint the reader with their unique properties. Chapter \ref{CH:Governing_equations_chapter} outlines the fundamental balance equations. These equations are then integrated with the extraction kinetic equation in Chapter \ref{CH: Extraction_model} to formulate the process model. Chapter \ref{CH: Parameter_estimation} introduces the maximum likelihood estimation technique, which is subsequently applied within the context of the process model. Chapter \ref{CH: Experiments} details the experimental procedures and the data acquisition necessary for estimating the kinetic parameters related to the extraction process. The study culminates in Chapters \ref{CH: Results} and \ref{CH: Conclusion}, where the parameter estimations and simulation outcomes are thoroughly analysed and discussed.

>>>>

	This study investigates the extraction of essential oil from chamomile flowers (Matricaria chamomilla L.) via supercritical fluid extraction techniques and modelling of this process. Chamomile is a medicinal herb widely cultivated in southern and eastern Europe—such as Germany, Hungary, France, and Russia. It can be found outside of Europe in Brazil as discussed by \citet{Singh2011}. This plant is distinguished by its hollow, bright gold cones, housing disc or tubular florets and surrounded by about fifteen white ray or ligulate florets. Chamomile has been used for its medicinal benefits, serving as an anti-inflammatory, antioxidant, mild astringent, and healing remedy. Chamomile's aqueous extract is widely used as a gentle sedative to calm nerves and mitigate anxiety, hysteria, nightmares, insomnia, and other sleep-related conditions, according to \citet{Srivastava2009}. \citet{Orav2010} reported that oil yields from dried chamomile samples ranged from 0.7 to 6.7 mL/kg. Notably, the highest yields of essential oil, between 6.1 and 6.7 mL/kg, were derived from chamomile sourced from Latvia and Ukraine, while chamomile from Armenia exhibited a lower oil content of 0.7 mL/kg.

@@

	%The organization of the study is as follows: Chapter \ref{CH: Thermodynamic} offers an overview of supercritical fluids, aiming to acquaint the reader with their unique properties. 
	Chapter \ref{CH:Governing_equations_chapter} outlines the fundamental balance equations. These equations are then integrated with the extraction kinetic equation in Chapter \ref{CH: Extraction_model} to formulate the process model. Chapter \ref{CH: Parameter_estimation} introduces the maximum likelihood estimation technique, which is subsequently applied within the context of the process model. Chapter \ref{CH: Experiments} details the experimental procedures and the data acquisition necessary for estimating the kinetic parameters related to the extraction process. The study culminates in Chapters \ref{CH: Results} and \ref{CH: Conclusion}, where the parameter estimations and simulation outcomes are thoroughly analysed and discussed.

### 2024-03-13 SFE_article_Parameter_Estimation_short - Chamomile/Short_Supercritical_fluid_extraction_of_essential_oil_from_chamomile.tex
13.03

\subsection{Supercritical fluids} \label{CH: Thermodynamic}
\subfile{Sections/Thermo_imp}

@@

The low Mach number condition leads to the incompressible condition: $\nabla \cdot \mathbf {u} =0$, which is valid for constant density (strict incompressible) or varying density flow. The restraint allows for the removal of acoustic waves, but also allows for large perturbations in density and/or temperature. The assumption is that the flow remains within a Mach number limit (normally less than 0.3) for any solution using such a constraint to be valid. In 1-D case the incompressibility condition becomes $\frac{du}{dz} = 0$, so the fluid velocity is constant.

@@

>>>>

%\subsection{Supercritical fluids} \label{CH: Thermodynamic}
%\subfile{Sections/Thermo_imp}

@@

The low Mach number condition leads to the incompressible condition: $\nabla \cdot u =0$, which is valid for constant density (strict incompressible) or varying density flow. The restraint allows for the removal of acoustic waves, but also allows for large perturbations in density and/or temperature. The assumption is that the flow remains within a Mach number limit (normally less than 0.3) for any solution using such a constraint to be valid. In 1-D case the incompressibility condition becomes $\frac{du}{dz} = 0$, so the fluid velocity is constant.

@@

\begin{figure*}[!h]
	\centering
	\begin{subfigure}{0.3\textwidth}
		\centering
		\includegraphics[trim = 0.0cm 0.0cm 0.0cm 0.0cm,clip, width=\textwidth]{/Results_estimation/Fit_Di_Gamma_1_4_correlation.png}
		\caption{Results of parameter estimation for experiments at $6.67\times 10^{-5}$ [kg/s] and temperature of 40 $[^\circ C]$}
		\label{fig: Fit_1_4_Di_Gamma_correlation}
	\end{subfigure}
	\hfill
	\begin{subfigure}{0.3\textwidth}
		\centering
		\includegraphics[trim = 0.0cm 0.0cm 0.0cm 0.0cm,clip, width=\textwidth]{/Results_estimation/Fit_Di_Gamma_5_8_correlation.png}
		\caption{Results of parameter estimation for experiments at $6.67\times 10^{-5}$ [kg/s] and temperature of 30 $[^\circ C]$}
		\label{fig: Fit_5_8_Di_Gamma_correlation}
	\end{subfigure}
	\hfill
	\begin{subfigure}{0.3\textwidth}
		\centering
		\includegraphics[trim = 0.0cm 0.0cm 0.0cm 0.0cm,clip, width=\textwidth]{/Results_estimation/Fit_Di_Gamma_9_12_correlation.png}
		\caption{Results of parameter estimation for experiments at $3.33\times 10^{-5}$ [kg/s]}
		\label{fig: Fit_9_12_Di_Gamma_correlation}
	\end{subfigure}
	\caption{Parameter estimation results}
	\label{fig: Fit_Di_Gamma_correlation}
\end{figure*}
